GPU-Accelerated Parallelization of Edge-Based Motion Detection Algorithms with the NVIDIA CUDA Framework

### 2015-12-01 documents/paper/paper.tex
Theory details for motion detection algorithm + images

\subsection{Non-Maximum Suppression and Selective Thresholding}
Given the gradient magnitude and direction of the Gaussian-filtered image, the final step of the edge detection algorithm is to determine which pixels should be selected as edge pixels, and which should be rejected. We define two procedures, \textit{non-maximum suppression} and \textit{selective thresholding}, to gain a reasonably accurate edge map from the gradient.
\par Non-maximum suppression selects and rejects edge pixels according to the following criteria:
\begin{enumerate}
	\item Pixels whose gradient magnitude is below a user-defined low threshold, $t_l$, is immediately rejected
	\item If the gradient magnitude of a pixel is greater than that of the pixels in either direction of the gradient angle of that pixel, then the pixel is accepted if its gradient magnitude is greater than a user-defined high threshold, $t_h$
\end{enumerate}
More precisely, a pixel at $(i, j)$ with gradient magnitude $G_{i, j} > t_l$ and angle $\theta$ is accepted if any of the following hold true:
\begin{itemize}
	\item $(-\frac{\pi}{6} < \theta < \frac{\pi}{6} \vee -\pi < \theta < -\frac{5 \pi}{6} \vee \frac{5\pi}{6} < \theta < \pi) \wedge G_{i, j} > G_{i, j \pm 1}$
	\item $(\frac{\pi}{6} < \theta < \frac{\pi}{3} \vee -\frac{5 \pi}{6} < \theta < -\frac{2 \pi}{3}) \wedge G_{i, j} > G_{i \pm 1, \pm j}$
	\item $(\frac{\pi}{3} < \theta < \frac{2 \pi}{3} \vee -\frac{2 \pi}{3} < \theta < -\frac{\pi}{3}) \wedge G_{i, j} > G_{i \pm 1, j}$
	\item $(\frac{2 \pi}{3} < \theta < \frac{5 \pi}{6} \vee -\frac{\pi}{3} < \theta < -\frac{\pi}{6}) \wedge G_{i, j} > G_{i \pm 1, \mp j}$
\end{itemize}
The selective thresholding technique extends on the procedure of non-maximum suppression. For any pixel that passes crtiteria (2) (maximized, relative to the neighboring pixels, along the gradient direction), consider an $\alpha$ x $\alpha$ box around the pixel. If any pixel within the box exceeds the low threshold $t_l$, then the pixel is accepted. This technique attempts to catch pixels that might ``connect" ridge pixels identified by criteria (2) but fail to satify the criterion themselves, in order to create more continuous edge lines.
\begin{algorithm}[h]
	\smaller
	\label{non-maximum-suppression-selective-thresholding-algorithm}
	\caption{Non-maximum suppression and selective thresholding: pseudocode for a serial, CPU implementation}

@@

\section{Motion Detection Algorithm}
Our motion detection algorithm compares the edges from two frames temporally separated by a small interval of time and attempts to estimate regions of the frame that exhibit movement or motion. This process is then repeated continuously to provide a real-time visualization of the regions of movement in a continuous video stream. The motion detection algorithm we present is composed of the following steps:
\begin{enumerate}
	\item Given two frames $\boldsymbol{F}_1$ and $\boldsymbol{F}_2$ separated by a time interval of $\Delta t$, compute its edges $\boldsymbol{E}_1$ and $\boldsymbol{E}_2$.
	\item Create a strict binary difference matrix $\boldsymbol{D}_0$ between $\boldsymbol{E}_1$ and $\boldsymbol{E}_2$ whose value is high at locations at which an edge is present in frame but not the other, and low at locations at which both frames have an edge or no edge.
	\item For each location $(i, j)$ in $\boldsymbol{D}_0$, a thresholded difference matrix $\boldsymbol{D}$ is created by considering a box of constant size $\beta$ x $\beta$ around that pixel and suppressing the value of the surrounding pixels to 0 if its value matches that at $\boldsymbol{D}_{0_{i, j}}$. This step is designed to reduce false positives of motion detection arising from stray movement of the frame (e.g. camera shake).
\end{enumerate}
More precisely, the matrix $\boldsymbol{D}$ referenced above is defined as
\[
\boldsymbol{D} = |\boldsymbol{E}_1 - \boldsymbol{E}_2| =
\begin{cases}
	255 & \boldsymbol{E}_{1_{i, j}} \neq \boldsymbol{E}_{2_{i, j}} \\
	0 & \boldsymbol{E}_{1_{i, j}} = \boldsymbol{E}_{2_{i, j}}
\end{cases}
\]
The computation of $\boldsymbol{D'}$ is determined by Algorithm \ref{difference-matrix-algorithm}.
\begin{algorithm}[h]
	\smaller
	\label{difference-matrix-algorithm}
	\caption{Determining the difference matrix between two frames, given their edge data: pseudocode for a serial, CPU implementation}

@@

\subsection{Thresholded Frame-by-Frame Difference Map}
Placeholder text

>>>>

\subsection{Non-Maximum Suppression and Selective Thresholding}
Given the gradient magnitude and direction of the Gaussian-filtered image, the final step of the edge detection algorithm is to determine which pixels should be selected as edge pixels, and which should be rejected. We define two procedures, \textit{non-maximum suppression} and \textit{selective thresholding}, to gain a reasonably accurate edge map from the gradient.
\par Non-maximum suppression selects and rejects edge pixels according to the following criteria:
\begin{enumerate}
	\item Pixels whose gradient magnitude is below a user-defined low threshold, $t_l$, is immediately rejected
	\item If the gradient magnitude of a pixel is greater than that of the pixels in either direction of the gradient angle of that pixel, then the pixel is accepted if its gradient magnitude is greater than a user-defined high threshold, $t_h$
\end{enumerate}
More precisely, a pixel at $(i, j)$ with gradient magnitude $G_{i, j} > t_l$ and angle $\theta$ is accepted if any of the following hold true:
\begin{itemize}
	\item $(-\frac{\pi}{6} < \theta < \frac{\pi}{6} \vee -\pi < \theta < -\frac{5 \pi}{6} \vee \frac{5\pi}{6} < \theta < \pi) \wedge G_{i, j} > G_{i, j \pm 1}$
	\item $(\frac{\pi}{6} < \theta < \frac{\pi}{3} \vee -\frac{5 \pi}{6} < \theta < -\frac{2 \pi}{3}) \wedge G_{i, j} > G_{i \pm 1, \pm j}$
	\item $(\frac{\pi}{3} < \theta < \frac{2 \pi}{3} \vee -\frac{2 \pi}{3} < \theta < -\frac{\pi}{3}) \wedge G_{i, j} > G_{i \pm 1, j}$
	\item $(\frac{2 \pi}{3} < \theta < \frac{5 \pi}{6} \vee -\frac{\pi}{3} < \theta < -\frac{\pi}{6}) \wedge G_{i, j} > G_{i \pm 1, \mp j}$
\end{itemize}
The selective thresholding technique extends on the procedure of non-maximum suppression. For any pixel that passes crtiteria (2) (maximized, relative to the neighboring pixels, along the gradient direction), consider an $\alpha$ x $\alpha$ box around the pixel. If any pixel within the box exceeds the low threshold $t_l$, then the pixel is accepted. This technique attempts to catch pixels that might ``connect" ridge pixels identified by criteria (2) but fail to satify the criterion themselves, in order to create more continuous edge lines.
\begin{figure}[]
	\centering
	\includegraphics[width=0.45\textwidth]{pcb_edges.jpg}
	\caption{Output of edge detection algorithm after non-maximum suppression and selective thresholding on the image of Figure \ref{pcb-original}, with parameters $t_l = 70$ and $t_h = 80$}
    \label{pcb-edges}
\end{figure}
\begin{algorithm}[h]
	\smaller
	\label{non-maximum-suppression-selective-thresholding-algorithm}
	\caption{Non-maximum suppression and selective thresholding: pseudocode for a serial, CPU implementation}

@@

\section{Motion Detection Algorithm}
Our motion detection algorithm compares the edges from two frames temporally separated by a small interval of time and attempts to estimate regions of the frame that exhibit movement or motion. This process is then repeated continuously to provide a real-time visualization of the regions of movement in a continuous video stream. The motion detection algorithm we present is composed of the following steps:
\begin{enumerate}
	\item Given two frames $\boldsymbol{F}_1$ and $\boldsymbol{F}_2$ separated by a time interval of $\Delta t$, compute its edges $\boldsymbol{E}_1$ and $\boldsymbol{E}_2$.
	\item Create a strict binary difference matrix $\boldsymbol{D}_0$ between $\boldsymbol{E}_1$ and $\boldsymbol{E}_2$ whose value is high at locations at which an edge is present in frame but not the other, and low at locations at which both frames have an edge or no edge.
	\item Apply a parameter-controlled movement tolerance thresholding operation to the matrix $\boldsymbol{D}_0$ to obtain matrix $\boldsymbol{D}$, in order to suppress potential false positives of detected motion from incidental camera movement.
\end{enumerate}
\par $\boldsymbol{F}_1$ and $\boldsymbol{F}_2$ are obtained by sampling frames from a continuous video stream, a task aided by the OpenCV C library. Thus, the time interval $\Delta t$ will be small enough for a real-time algorithm. The edges $\boldsymbol{E}_1$ and $\boldsymbol{E}_2$ are computed with the edge detection algorithm of Section II. Below, we explore, in greater depth, the procedures for building a thresholded diffrence matrix and finally estimating the motion area by analyzing a spatial difference density map.

\subsection{Thresholded Frame-by-Frame Difference Map}
The preliminary requirement in determining areas of motion from two frames is a systematic method of determining and quantifying the differences between two frames separated by a small $\Delta t$. Intuitively, areas of motion within the time interval $\Delta t$ should exist at locations at which the difference between the two frames is large. This process is then continuously repeated for every pair of input frames in order to create a real-time motion detector. In our algorithm, we quantify differences on a binary basis using edge data from our edge detection algorithm, then determine motion locations from a thresholded density map of the difference matrix $\boldsymbol{D}$.
\par More formally, we begin with a matrix $\boldsymbol{D}_0$ created from a simple comparison of $\boldsymbol{E}_1$ and $\boldsymbol{E}_2$.
\[
\boldsymbol{D}_0 = |\boldsymbol{E}_1 - \boldsymbol{E}_2| =
\begin{cases}
	255 & \boldsymbol{E}_{1_{i, j}} \neq \boldsymbol{E}_{2_{i, j}} \\
	0 & \boldsymbol{E}_{1_{i, j}} = \boldsymbol{E}_{2_{i, j}}
\end{cases}
\]
Then, for each location $(i, j)$ in $\boldsymbol{D}_0$, a thresholded difference matrix $\boldsymbol{D}$ is created by considering a box of constant size $\beta$ x $\beta$ around that pixel and suppressing the value of the surrounding pixels to 0 if its value matches that at $\boldsymbol{D}_{0_{i, j}}$. This step is designed to reduce false positives of motion detection arising from stray movement of the frame (e.g. camera shake). The computation of $\boldsymbol{D}$ is demonstrated in Algorithm \ref{difference-matrix-algorithm}. The output matrix is then used as input to building a spatial difference density map, from which we estimate motion locations.
\begin{figure}[H]
	\centering
	\includegraphics[width=0.45\textwidth]{desk_1.jpg}
	\caption{Example of difference matrix computation: original initial frame $\boldsymbol{F}_1$ of dimensions $5312$ x $2988$}
    \label{desk-1-original}
\end{figure}
\begin{figure}[H]
	\centering
	\includegraphics[width=0.45\textwidth]{desk_2.jpg}
	\caption{Example of difference matrix computation: original initial frame $\boldsymbol{F}_2$ (the item displaced between the two frames is the highlighter at the center)}
    \label{desk-2-original}
\end{figure}
\begin{figure}[H]
	\centering
	\includegraphics[width=0.45\textwidth]{desk_edges_1.jpg}
	\caption{Example of difference matrix computation: edges of the initial frame $\boldsymbol{E}_1$ computed with parameters $k = 5$, $\sigma = 5$, $t_l = 15$, $t_h = 25$}
    \label{desk-1-edges}
\end{figure}
\begin{figure}[H]
	\centering
	\includegraphics[width=0.45\textwidth]{desk_edges_2.jpg}
	\caption{Example of difference matrix computation: edges of the initial frame $\boldsymbol{E}_2$ computed with parameters $k = 5$, $\sigma = 5$, $t_l = 15$, $t_h = 25$}
    \label{desk-2-edges}
\end{figure}
\begin{figure}[H]
	\centering
	\includegraphics[width=0.45\textwidth]{desk_difference.jpg}
	\caption{Example of difference matrix computation: difference matrix $\boldsymbol{D}$ with parameter $\beta = 12$}
    \label{desk-difference}
\end{figure}
\begin{algorithm}[]
	\smaller
	\label{difference-matrix-algorithm}
	\caption{Determining the difference matrix between two frames, given their edge data: pseudocode for a serial, CPU implementation}

@@
